%!TEX root = ../../main.tex

\chapter*{Selbständigkeitserklärung gemäß § 60 Abs. 8 AllgStuPO}


Hiermit erkläre ich, dass ich die vorliegende Arbeit selbstständig und eigenhändig sowie ohne unerlaubte Hilfe und ausschließlich unter Verwendung der aufgeführten Quellen und Hilfsmittel angefertigt habe.  
Die selbstständige und eigenhändige Anfertigung versichere ich an Eides statt. Hinweis zum Einsatz von KI-Werkzeugen:
Das KI-Tool ChatGPT wurde ausschließlich zur grammatikalischen Überprüfung und Verbesserung des Textes verwendet.
Die inhaltliche Ausarbeitung stammt von mir. Die Satzung zur Sicherung guter wissenschaftlicher Praxis an der TU Berlin vom 15. Februar 2023.
\url{https://www.static.tu.berlin/fileadmin/www/10002457/K3-AMBl/Amtsblatt\_2023/Amtliches\_Mitteilungsblatt\_Nr.\_16\_vom\_30.05.2023.pdf} habe ich zur Kenntnis
genommen.
Ich erkläre weiterhin, dass ich die Arbeit in gleicher oder ähnlicher Form noch keiner anderen Prüfungsbehörde
vorgelegt habe.


\vspace{3cm}

\noindent\begin{tabular}{ll}
    \makebox[65mm]{\hrulefill} & \makebox[65mm]{\hrulefill} \\
    Datum, Ort & Unterschrift der/des Studierenden
\end{tabular}

\chapter*{Agreement on utilization rights}
{

\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

According to § 60 of the AllgStuPO, students of the TU Berlin must complete a final thesis during their studies. The totality of the achievements associated with a final thesis are based not only on the written documentation and the commitment of the student, but also on the supervisory effort of the department, including the preparation of the assignment, the specific requirements, guidance of the students, help and advice in organizational, technical and scientific terms. Therefore, the student is the sole author of the written work, but not the sole author of the entirety of the services associated with a thesis.


By supervising the student, the Technische Universität Berlin, represented by the Chair of Astronautics, becomes co-author of the entirety of the achievements and the associated results according to § 8 section 1 of the German Copyright Act (Urheberrechtsgesetz).


For non-profit use in teaching and research, the Chair may use the results of the present work. As co-author, the chair is entitled to do so according to § 8 section 2 of the Copyright Act. If the work is written independently in such a way that there is no co-authorship, this right of use is granted by the student according to § 31 section 2 of the Copyright Act. This right of use is unrestricted and includes content of any kind (e.g. documentation, presentations, photos, videos, procedures, drafts, drawings, software including source code and the like) with naming of the author.


Due to the reference to ongoing research projects, publication without explicit permission of the department will not be permitted to the student. This is to be checked and approved by the department in each individual case.


Any commercial use of either site will only occur with the consent of all authors of the present work, with appropriate participation in the earnings.

}
\vspace{3cm}

\noindent\begin{tabular}{ll}
    \makebox[65mm]{\hrulefill} & \makebox[65mm]{\hrulefill} \\
    Date, Place & Signature of the Student
\end{tabular}